---
title: "Problem Section 3"
subtitle:  "Mon Oct 16, 2023"
graphics: yes
output: 
        pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * * 
\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence.
The specific tasks include: 

   - Identifying the range of discrete and continuous random variables
   
   - Deriving the PMF and CDF of a discrete random variable
   
   - Recognizing properties of CDFs
   
   - Back up and support work with relevant explanations

\end{shaded}


* * *

### Exercises 


1. For each situation, give the range of the random variable $X$. Then state whether the random variable is discrete or continuous.

a. Three fair coins are tossed independently until they show the same side. $X$ is the number of tosses needed.

The range is $\{1, 2, 3, \dots\}$. Another interpretation: $\{3, 6, 9, \dots \}$. Since $X$ can only take a countable number of values, it is discrete.

b. Using modern aptitude testing and screening procedures, Rote Industries has greatly reduced its employee turnover. Recent records show that 90% of the company’s new assembly line workers stay with the company for at least a year. $X$ is the number of employees among 12 recently hired that will still be with the company in another year.

The range is $\{0, 1, 2, \dots, 12 \}$. Again, for the same reason as part (a), $X$ is discrete.
		
c. In a busy 24 hour copy center, minor malfunctions of the copy machine occur randomly. $X$ denotes the time (in hours) until the first minor malfunction occurs.

The range is $[0, \infty)$. Here, $X$ can take any value in an interval and can (in theory) be measured with infinite precision, so it is a continuous random variable. 
	 
2. Suppose you toss a fair coin 3 times. The sample space for this experiment is shown below:
$$S = \{ HHH, HHT, HTH, HTT, TTT,TTH, THT, THH \}.$$

a. Write the PMF of $X$ defined as the difference between number of heads and number of tails.

Consider the cases:

\begin{enumerate}
\item 0 head, 3 tails (heads - tails = -3)
\item 1 head, 2 tails  (heads - tails = -1)
\item 2 head, 1 tails (heads - tails = 1)
\item 3 heads, 0 tails (heads - tails = 3)
\end{enumerate}

So the PMF is

\begin{tabular}{|c|c|c|c|c|} \hline
$x$ & -3 & -1 & 1 & 3 \\ \hline
$f(x)$ &  1/8 & 3/8 & 3/8 & 1/8 \\ \hline
\end{tabular}


b. Give the CDF of $X$.

\begin{align*}
F(x) &= P(X \leq x) \\
&= \left \{
\begin{array}{cc}
0 & \text{ if } x < -3 \\ 
\frac{1}{8} & \text{ if } -3 \leq x < -1 \\ 
\frac{4}{8} & \text{ if } -1 \leq x < 1 \\ 
\frac{7}{8} & \text{ if } 1 \leq x < 3 \\ 
1 & \text{ if } x \geq 3
\end{array} 
\right.
\end{align*}

c. Draw the probability histogram of $X$ in R.

```{r}
#be sure to add library(tidyverse) in setup chunk. 

ht_df <- data.frame(x = c(-3,-1,1,3), f = c(1/8,3/8,3/8,1/8))

ggplot(data = ht_df, mapping = aes(x=x,y=f)) +
  geom_col(width=1, alpha=0.5, fill="gold",color="purple") +
  geom_text(mapping=aes(label = round(f,4),y = f + 0.01)) +
  labs(title="Probability Histogram of X", y = "Probability of X", x = "X=x") +
  scale_x_continuous(breaks=c(-3,-1,1,3))

```


	  
3. Which of the following is a valid CDF? (choose one; also point out what's wrong with the ones you didn't choose)

a. 

\begin{align*} 
        F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x \leq 3 \\
                 \frac{3}{4} & 3 <  x < 5 \\
                 1 & 5 \leq x \end{array} \right. \end{align*}


This is not a valid CDF because it is not right continuous at $x=3$. 

b.

\begin{align*} 
        F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x < 3 \\
                 \frac{3}{4} & 3 \leq x < 5 \\
                 1 & 5 \leq x \end{array} \right. \end{align*}


This is a valid CDF because it is non-decreasing and right continuous. 

c. 

\begin{align*} 
      F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{3}{4} & 0 \leq x < 3 \\
                 \frac{1}{4} & 3 \leq x < 5 \\
                 1 & 5 \leq x \end{array} \right. \end{align*}


This is not a valid CDF becaue it decreases at $x=3$.

d.

\begin{align*} 
       F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x < 3 \\
                 \frac{3}{4} & 3 \leq x < 5 \\
                 \frac{7}{8} & 5 \leq x \end{array}   
                 \right.                      
\end{align*}  



This is not a valid CDF since it does not converge to 1  as $x \rightarrow \infty$.
